Considering scale in multilayer animal social networks

### 2020-06-20 paper/manuscript.Rmd
no rangifer

<!-- Figure: metrics by social scale -->
(ref:socres) Varying scale in social distance threshold. For each social distance threshold (100 - 1000 m by steps of 100 m), multilayer networks were constructed of caribou social associations (*R. tarandus*, n = `r DT[, uniqueN(get(idcol))]` females) on Fogo Island, Newfoundland between `r DT[, format(min(datetime), '%F')]` and `r DT[, format(max(datetime), '%F')]` across three habitat classes (forage, forest, and open). Graph strength (mean of individuals in bold) showed a sharp increase between 5 m and 100 m for all habitat classes. After 100 m, graph strength continued to increase at a slower rate. Open and forage layers showed higher variation in graph strength between individuals while the forest layer had less variation and lower mean graph strength.

<!-- Figure: metrics by land cover resolution -->
(ref:lcres) Varying spatial scale of land cover raster. For each land cover resolution (5, 25, 50, 75, 100, 250 and 500 m), multilayer networks were constructed of caribou social associations (*R. tarandus*, n = `r DT[, uniqueN(get(idcol))]` females) on Fogo Island, Newfoundland between `r DT[, format(min(datetime), '%F')]` and `r DT[, format(max(datetime), '%F')]` across three habitat classes (forage, forest, and open). A, B, C) Three land cover rasters: the original resolution (30 m) and two aggregated rasters (500 m and 1000 m). D) Edge overlap of habitat layers across spatial resolutions. Open and forest layers show consistent edge overlap with increasing spatial resolution while forage shows a sharp decline in edge overlap between 30 m and 600 m. Above 600 m, the forage layers are stable. E) Graph strength (mean of individuals in bold) within habitat layers across spatial resolution. Forage layers showed high variation in graph strength with many individuals dropping to 0 after 500 m and others increasing with spatial resolution. Open layers and forest layers were relatively stable across spatial resolutions, with higher mean graph strength in open compared to forest.

<!-- Figure: metrics by number of observations -->
(ref:nobs) Varying scale in number of observations. Multilayer networks were generated using a  sample number of observations (10 - 1000 observations by steps of 10) of caribou social associations (*R. tarandus*, n = `r DT[, uniqueN(get(idcol))]` females) on Fogo Island, Newfoundland between `r DT[, format(min(datetime), '%F')]` and `r DT[, format(max(datetime), '%F')]` across three habitat classes (forage, forest, and open). A) Graph strength was highly variable at low number of observations (less than 100) for all habitat layers and relatively stable after 200 observations. B) Edge overlap increased for all habitat layers with increasing number of observations. Open and forage layers had higher edge overlap than forest layers. C) Multidegree increased for all individuals with increasing number of observations. Individuals showed high variability (`r nobs[nobs == max(nobs), paste(range(multideg), collapse = ' - ')]`) at the maximal number of observations (1000).

<!-- Figure: Time window -->
(ref:temp) Varying temporal scale in multilayer networks. Multilayer networks were generated of caribou social associations (*R. tarandus*, n = `r DT[, uniqueN(get(idcol))]` females) on Fogo Island, Newfoundland within ordinal time windows (20 ordinal sample periods of 35-36 days) between `r DT[, format(min(datetime), '%F')]` and `r DT[, format(max(datetime), '%F')]`. A) Edge overlap across time windows showed two periods of low overlap (time windows 2-5 and 12-15), and three periods of high overlap (time windows 1, 6-11, 15-20). B) Ordinal network layers showing unweighted edges between individuals within time windows. Like edge overlap, individuals were more connected in time windows 1, 6-11, and 15-20, compared to 2-5 and 12-15. C) Graph strength within time windows showing individuals as thin grey lines and mean across individuals as thick grey lines. Graph strength peaked in time window 9, with noticeable higher strength than the following year.

@@

The emergent relationship between habitat and sociality has been explored by the
distribution and phenology of resources and use of social information. For
example, cougar (*Puma concolor*) home ranges overlapped more in regions with
high hunting opportunity, a proxy for resource richness [@Elbroch_2015]. Caribou
(*R. tarandus*) social network graph strength differed between seasons, where 
graph strength was stronger in winter and contrastingly low in summer when
resource availability are more homogeneous [@Peignier_2019]. Bison (*Bison
bison*) showed stronger habitat selection in larger groups, and avoidance of
predation risk was a main factor contributing to their habitat selection
[@Fortin_2009]. These examples, comparing sociality and habitat selection across
temporal or spatial contexts, are implicitly multilayer analyses.

@@

(ref:ml) Visualization of multilayer network representing social association of caribou (*Rangifer tarandus*) on Fogo Island, Newfoundland between `r DT[, format(min(datetime), '%F')]` and `r DT[, format(max(datetime), '%F')]`. Individual caribou (n = `r DT[, uniqueN(get(idcol))]` females) are show as nodes and each individual is represented by a unique color. Twelve layers represent combinations of four time windows and three habitat classes (open, forest and forage). Intralayer edges connect pairs of individuals that associated according to spatial and temporal thresholds described above, and are scaled in line thickness according to strength of association (SRI).

>>>>

<!-- Figure: metrics by social scale -->
(ref:socres) Varying scale in social distance threshold. For each social distance threshold (100 - 1000 m by steps of 100 m), multilayer networks were constructed of caribou social associations (n = `r DT[, uniqueN(get(idcol))]` females) on Fogo Island, Newfoundland between `r DT[, format(min(datetime), '%F')]` and `r DT[, format(max(datetime), '%F')]` across three habitat classes (forage, forest, and open). Graph strength (mean of individuals in bold) showed a sharp increase between 5 m and 100 m for all habitat classes. After 100 m, graph strength continued to increase at a slower rate. Open and forage layers showed higher variation in graph strength between individuals while the forest layer had less variation and lower mean graph strength.

<!-- Figure: metrics by land cover resolution -->
(ref:lcres) Varying spatial scale of land cover raster. For each land cover resolution (5, 25, 50, 75, 100, 250 and 500 m), multilayer networks were constructed of caribou social associations (n = `r DT[, uniqueN(get(idcol))]` females) on Fogo Island, Newfoundland between `r DT[, format(min(datetime), '%F')]` and `r DT[, format(max(datetime), '%F')]` across three habitat classes (forage, forest, and open). A, B, C) Three land cover rasters: the original resolution (30 m) and two aggregated rasters (500 m and 1000 m). D) Edge overlap of habitat layers across spatial resolutions. Open and forest layers show consistent edge overlap with increasing spatial resolution while forage shows a sharp decline in edge overlap between 30 m and 600 m. Above 600 m, the forage layers are stable. E) Graph strength (mean of individuals in bold) within habitat layers across spatial resolution. Forage layers showed high variation in graph strength with many individuals dropping to 0 after 500 m and others increasing with spatial resolution. Open layers and forest layers were relatively stable across spatial resolutions, with higher mean graph strength in open compared to forest.

<!-- Figure: metrics by number of observations -->
(ref:nobs) Varying scale in number of observations. Multilayer networks were generated using a  sample number of observations (10 - 1000 observations by steps of 10) of caribou social associations (n = `r DT[, uniqueN(get(idcol))]` females) on Fogo Island, Newfoundland between `r DT[, format(min(datetime), '%F')]` and `r DT[, format(max(datetime), '%F')]` across three habitat classes (forage, forest, and open). A) Graph strength was highly variable at low number of observations (less than 100) for all habitat layers and relatively stable after 200 observations. B) Edge overlap increased for all habitat layers with increasing number of observations. Open and forage layers had higher edge overlap than forest layers. C) Multidegree increased for all individuals with increasing number of observations. Individuals showed high variability (`r nobs[nobs == max(nobs), paste(range(multideg), collapse = ' - ')]`) at the maximal number of observations (1000).

<!-- Figure: Time window -->
(ref:temp) Varying temporal scale in multilayer networks. Multilayer networks were generated of caribou social associations (n = `r DT[, uniqueN(get(idcol))]` females) on Fogo Island, Newfoundland within ordinal time windows (20 ordinal sample periods of 35-36 days) between `r DT[, format(min(datetime), '%F')]` and `r DT[, format(max(datetime), '%F')]`. A) Edge overlap across time windows showed two periods of low overlap (time windows 2-5 and 12-15), and three periods of high overlap (time windows 1, 6-11, 15-20). B) Ordinal network layers showing unweighted edges between individuals within time windows. Like edge overlap, individuals were more connected in time windows 1, 6-11, and 15-20, compared to 2-5 and 12-15. C) Graph strength within time windows showing individuals as thin grey lines and mean across individuals as thick grey lines. Graph strength peaked in time window 9, with noticeable higher strength than the following year.

@@

The emergent relationship between habitat and sociality has been explored by the
distribution and phenology of resources and use of social information. For
example, cougar (*Puma concolor*) home ranges overlapped more in regions with
high hunting opportunity, a proxy for resource richness [@Elbroch_2015]. Caribou
social network graph strength differed between seasons, where 
graph strength was stronger in winter and contrastingly low in summer when
resource availability are more homogeneous [@Peignier_2019]. Bison (*Bison
bison*) showed stronger habitat selection in larger groups, and avoidance of
predation risk was a main factor contributing to their habitat selection
[@Fortin_2009]. These examples, comparing sociality and habitat selection across
temporal or spatial contexts, are implicitly multilayer analyses.

@@

(ref:ml) Visualization of multilayer network representing social association of caribou on Fogo Island, Newfoundland between `r DT[, format(min(datetime), '%F')]` and `r DT[, format(max(datetime), '%F')]`. Individual caribou (n = `r DT[, uniqueN(get(idcol))]` females) are show as nodes and each individual is represented by a unique color. Twelve layers represent combinations of four time windows and three habitat classes (open, forest and forage). Intralayer edges connect pairs of individuals that associated according to spatial and temporal thresholds described above, and are scaled in line thickness according to strength of association (SRI).
